---
title: "Lab 04 / Exam 2 Rough Draft"
subtitle: "Hypothesis tests and confidence intervals"
author: "YOUR NAME HERE"
format:
  html:
    self-contained: true
    toc: true
    toc_float: true
    number_section: false
    highlight: tango
editor: visual
editor_options:
  chunk_output_type: console
---

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(infer)

gss2016 <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSXStAOCL3Y19OuZAoREVDHKzgiBycVo5h5NRvkU3l4IB81kx-BMU4hurcpWSdfYszJIRQJZ7QblJ8-/pub?output=csv")

beach_vb <- readRDS("beach_vb_lab04.RDS")
```

## Exercise 1

a\. YOUR ANSWER HERE

b\. $paste your LaTeX syntax inside these dollar signs$

c\. $paste your LaTeX syntax inside these dollar signs$

## Exercise 2

TYPE YOUR LaTeX syntax here - make sure to put it inside double dollar signs to center the equations.

## Exercise 3

```{r}
# part a code 

```

b\. YOUR ANSWER HERE

c\. YOUR ANSWER HERE

d\. YOUR ANSWER HERE

## Exercise 4

```{r}
# part a code 

```

b\. YOUR ANSWER HERE

## Exercise 5

a\. YOUR ANSWER HERE

b\. YOUR ANSWER HERE

```{r}
# part c code 

```

d\. YOUR ANSWER HERE

e\. YOUR ANSWER HERE

f\. YOUR ANSWER HERE

g\. YOUR ANSWER HERE

## Exercise 6

```{r}
# part a code 


```

b\. YOUR ANSWER HERE

c\. YOUR ANSWER HERE

## Exercise 7

a\. YOUR ANSWER HERE

b\. YOUR ANSWER HERE

c\. YOUR ANSWER HERE

## Exercise 8

```{r}
# part a code 


```

b\. YOUR ANSWER HERE

c\. YOUR ANSWER HERE

## Exercise 9

```{r}
# part a code 


```

b\. YOUR ANSWER HERE

c\. YOUR ANSWER HERE

d\. YOUR ANSWER HERE
