---
title: "Lab 05 slides"
subtitle: "MATH 130"
author: "Katie Fitzgerald"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "slides.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(gghighlight)
library(broom)
library(palmerpenguins)
```



class: middle

# Regression in R (new code for Lab 05)

---

## The data

```{r, warning = FALSE}
scorecard <- read_csv("data/scorecard.csv")
glimpse(scorecard)
```


---

## Fitting a regression model

```{r, eval = FALSE}
model_name <- lm(y_variable ~ x_variable, dataset)

tidy(model_name)
```

+ `lm` stands for "linear model" and is the function that actually fits the least squares regression model

--

+ Nothing will appear in output after running the first line of code, but a new object called `model_name` will appear in your Environment

--

+ `tidy()` is the function we use to see the regression table output

--

+ Whatever you name your model is what should go inside `tidy()`

---

## Fitting a regression model: `lm()` and `tidy()`

```{r fit-tidy, fig.show='hide'}
model1 <- lm(pell_pct ~ firstgen_pct, scorecard)

tidy(model1)
```



--

The `estimate` column contains the intercept $(b_0)$ and slope $(b_1)$!

When you see a table of output like this, you don't have to calculate the intercept or slope!

--

$$\hat{y} = 6.92 + 0.967x$$
---

## Visualizing a regression model

.panelset[
.panel[.panel-name[Output]
```{r ref.label = "reg-viz", echo = FALSE, warning = FALSE, out.width = "70%", fig.width = 8}

```
]
.panel[.panel-name[Code]

```{r reg-viz, fig.show='hide', eval = FALSE}
ggplot(scorecard, aes(x = firstgen_pct, y = pell_pct)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```
]
]

---

## Calculating $R^2$

The `glance()` function will calculate $R^2$

```{r}
glance(model1)
glance(model1)[1]
```

Whatever you name your model is what you should put inside `glance()`

---

## Recap: Regression in R

```{r, eval = FALSE}
model_name <- lm(______ ~ ______, ______)

tidy(______)

ggplot(______, aes(x = ______, y = ______)) +
  geom____() +
  geom____(method = "lm", se = FALSE)

glance(______)[1]
```

--

```{r, eval = FALSE}
model_name <- lm(y_variable ~ x_variable, dataset)

tidy(model_name)

ggplot(dataset, aes(x = x_variable, y = y_variable)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

glance(model_name)[1]
```

