---
title: "Lab 03 slides"
subtitle: "MATH 130"
author: "Katie Fitzgerald"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "slides.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(gghighlight)
library(broom)
library(palmerpenguins)
```

class: middle

# Recap: `ggplot()`, `summarize()`, `group_by()`, `summary()`

---

## `ggplot()`

Structure of the code for plots can be summarized as

```{r eval = FALSE}
ggplot(dataset, aes(x = [x-variable], 
                    y = [y-variable])) +
   geom_xxx()
```

Examples: (see Labs 01 & 02 for more)

```{r, eval = FALSE}
ggplot(penguins, aes(x = body_mass_g, y = species)) +
  geom_boxplot()

ggplot(basketball, aes(x = age)) + 
  geom_histogram(binwidth = 2, color = "white")

ggplot(penguins, aes(x = bill_length_mm, 
                     y = flipper_length_mm)) +
  geom_point()
```

---

## `summarize()`

General structure of `summarize()`, used to compute summary statistics:

```{r, eval = FALSE}
dataset %>% 
  summarize(user_defined_label = function(variable, na.rm = TRUE))
```

Example: 

```{r}
penguins %>% 
  summarize(avg_body_mass = mean(body_mass_g, na.rm = TRUE))
```


---

# One summary statistic at a time

.panelset[
.panel[.panel-name[Output]
```{r ref.label = "one-stat", echo = FALSE, warning = FALSE, out.width = "70%", fig.width = 8}

```
]
.panel[.panel-name[Code]

```{r one-stat, fig.show='hide', eval = FALSE}
penguins %>% 
  summarize(avg_body_mass = mean(body_mass_g, na.rm = TRUE))

penguins %>% 
  summarize(sd_body_mass = sd(body_mass_g, na.rm = TRUE))
```
]
]

---

# Two summary statistics at a time

.panelset[
.panel[.panel-name[Output]
```{r ref.label = "two-stats", echo = FALSE, warning = FALSE, out.width = "70%", fig.width = 8}

```
]
.panel[.panel-name[Code]

```{r two-stats, fig.show='hide', eval = FALSE}
penguins %>% 
  summarize(avg_body_mass = mean(body_mass_g, na.rm = TRUE),
            sd_body_mass = sd(body_mass_g, na.rm = TRUE))
```
]
]

---

## `group_by()`

Add `group_by()` before `summarize()` to calculate summary statistics **for each group** (i.e. for each level of a categorical variable)


General structure: 

```{r, eval = FALSE}
dataset %>% 
  group_by(categorical_variable) %>% 
  summarize(user_defined_label = function(variable, na.rm = TRUE))
```

---

## Summary statistics for multiple groups

.panelset[
.panel[.panel-name[Output]
```{r ref.label = "group-by", echo = FALSE, warning = FALSE, out.width = "70%", fig.width = 8}

```
]
.panel[.panel-name[Code]

```{r group-by, fig.show='hide', eval = FALSE}
penguins %>% 
  group_by(species) %>% 
  summarize(avg_body_mass = mean(body_mass_g, na.rm = TRUE),
            sd_body_mass = sd(body_mass_g, na.rm = TRUE))
```
]
]

---

## `summary()` and `$`

Use the dollar sign to access an individual variable in a dataset

General structure:

```{r, eval = FALSE}
summary(dataset$variable)
```

Example: 

```{r}
summary(penguins$body_mass_g)
```


---

class: middle

# Regression in R (new code for Lab 03)

---

## The data

```{r, warning = FALSE}
scorecard <- read_csv("data/scorecard.csv")
glimpse(scorecard)
```


---

## Fitting a regression model

```{r, eval = FALSE}
model_name <- lm(y_variable ~ x_variable, dataset)

tidy(model_name)
```

+ `lm` stands for "linear model" and is the function that actually fits the least squares regression model

--

+ Nothing will appear in output after running the first line of code, but a new object called `model_name` will appear in your Environment

--

+ `tidy()` is the function we use to see the regression table output

--

+ Whatever you name your model is what should go inside `tidy()`

---

## Fitting a regression model: `lm()` and `tidy()`

```{r fit-tidy, fig.show='hide'}
model1 <- lm(pell_pct ~ firstgen_pct, scorecard)

tidy(model1)
```



--

The `estimate` column contains the intercept $(b_0)$ and slope $(b_1)$!

When you see a table of output like this, you don't have to calculate the intercept or slope!

--

$$\hat{y} = 6.92 + 0.967x$$
---

## Visualizing a regression model

.panelset[
.panel[.panel-name[Output]
```{r ref.label = "reg-viz", echo = FALSE, warning = FALSE, out.width = "70%", fig.width = 8}

```
]
.panel[.panel-name[Code]

```{r reg-viz, fig.show='hide', eval = FALSE}
ggplot(scorecard, aes(x = firstgen_pct, y = pell_pct)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```
]
]

---

## Calculating $R^2$

The `glance()` function will calculate $R^2$

```{r}
glance(model1)
glance(model1)[1]
```

Whatever you name your model is what you should put inside `glance()`

---

## Recap: Regression in R

```{r, eval = FALSE}
model_name <- lm(______ ~ ______, ______)

tidy(______)

ggplot(______, aes(x = ______, y = ______)) +
  geom____() +
  geom____(method = "lm", se = FALSE)

glance(______)[1]
```

--

```{r, eval = FALSE}
model_name <- lm(y_variable ~ x_variable, dataset)

tidy(model_name)

ggplot(dataset, aes(x = x_variable, y = y_variable)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

glance(model_name)[1]
```

