---
title: "Lab 04 slides"
subtitle: "MATH 130"
author: "Katie Fitzgerald"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "slides.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggdist)
library(distributional)
```

class: middle

# Normal distribution in R: `pnorm()` and `qnorm()`

---

# `pnorm()`

.pull-left[
Recall: Z-scores follow a normal distribution with mean 0 and standard deviation 1

To calculate $P(Z < 1.7)$:

```{r}
pnorm(1.7, mean = 0, sd = 1)
```

]

.pull-right[

```{r, echo = FALSE, warning = FALSE, message = FALSE}
ggplot() +
  stat_dist_halfeye(aes(y = "N(0,1)", 
                        dist = dist_normal(0, 1),
                        fill = stat(x < 1.7))) +
  labs(y = "", x = "x") +
  #viridis color scale is color-blind friendly
  #set direction = -1 so TRUE is dark instead of light
  scale_fill_viridis_d(direction = -1, guide = "none") +
  coord_cartesian(expand = FALSE, ylim = c(0.9, 2)) +
  theme_minimal() +
  theme(text = element_text(size = 40)) 
```

]

--

By default, R assumes `mean = 0`, and `sd = 1`, so we can leave these arguments off if our context matches the default:

```{r}
pnorm(1.7)
```

---

# "Upper tail" probabilities (yellow)

.pull-left[
By default, R will tell you the LOWER tail probability. In other words, it tells you the probability that falls to the LEFT of the cutoff (purple shaded region).

$P(Z < -0.5)$

```{r}
pnorm(-0.5)
```

]

.pull-right[

```{r, echo = FALSE, warning = FALSE, message = FALSE}
ggplot() +
  stat_dist_halfeye(aes(y = "N(0,1)", 
                        dist = dist_normal(0, 1),
                        fill = stat(x < -0.5))) +
  labs(y = "", x = "x") +
  #viridis color scale is color-blind friendly
  #set direction = -1 so TRUE is dark instead of light
  scale_fill_viridis_d(direction = -1, guide = "none") +
  coord_cartesian(expand = FALSE, ylim = c(0.9, 2)) +
  theme_minimal() +
  theme(text = element_text(size = 40)) 
```

If we want the UPPER (right) tail probability (yellow shaded region), we have to add `lower.tail = FALSE`:

$P(Z > -0.5)$

```{r}
pnorm(-0.5, lower.tail = FALSE)
```

]
---

# "Upper tail" probabilities (yellow)

.pull-left[

Notice, $P(Z > -0.5) = 1 - P(Z < -0.5)$. Or, in other words, the yellow shaded region is 1 minus the purple shaded region. So the following are two equivalent ways of calculating the upper tail probability:

```{r}
pnorm(-0.5, lower.tail = FALSE)
1 - pnorm(-0.5)
```

]

.pull-right[

```{r, echo = FALSE, warning = FALSE, message = FALSE}
ggplot() +
  stat_dist_halfeye(aes(y = "N(0,1)", 
                        dist = dist_normal(0, 1),
                        fill = stat(x < -0.5))) +
  labs(y = "", x = "x") +
  #viridis color scale is color-blind friendly
  #set direction = -1 so TRUE is dark instead of light
  scale_fill_viridis_d(direction = -1, guide = "none") +
  coord_cartesian(expand = FALSE, ylim = c(0.9, 2)) +
  theme_minimal() +
  theme(text = element_text(size = 40)) 
```

]


---

# `pnorm()` vs. `qnorm()`

+ `pnorm()` is for when we know the cutoff value (e.g. 1.7, or -0.5), and we want to know the probability (i.e. how much is shaded).

+ `qnorm()` is for when we know the probability (i.e. how much is shaded), and we want to know the cutoff value.

```{r}
pnorm(1.7)

qnorm(0.9554345)
```

+ They are INVERSE functions of one another

---

# `qnorm()`

`qnorm()` is good for finding percentiles. 

For example, what is the 75th percentile of SAT scores, which follow a normal distribution with mean 1500 and sd 300?

.pull-left[

```{r}
qnorm(0.75, mean = 1500, sd = 300)
```

]

.pull-right[

```{r, echo = FALSE, warning = FALSE, message = FALSE}
ggplot() +
  stat_dist_halfeye(aes(y = "N(0,1)", 
                        dist = dist_normal(1500, 300),
                        fill = stat(x < qnorm(.75, 1500, 300)))) +
  labs(y = "", x = "x") +
  #viridis color scale is color-blind friendly
  #set direction = -1 so TRUE is dark instead of light
  scale_fill_viridis_d(direction = -1, guide = "none") +
  coord_cartesian(expand = FALSE, ylim = c(0.9, 2)) +
  theme_minimal() +
  theme(text = element_text(size = 40)) 
```

]

What do you expect the following code to output?

```{r, eval = FALSE}
qnorm(0.5)
```


