---
title: "NYPD Stop, Question, Frisk"
subtitle: "Lab 02"
author: "YOUR NAME HERE"
format:
  html:
    self-contained: true
    toc: true
    toc_float: true
    number_section: false
    highlight: tango
editor: visual
editor_options:
  chunk_output_type: console
---

# Getting started: NYPD data

```{r, message = FALSE}
library(tidyverse) #load package

sqf_2023 <- read_rds("sqf_2023.rds") #loads data
```

```{r}
#creates dataset with fewer columns needed for analysis
sqf_selected <- sqf_2023 |> 
  select(1:15)

glimpse(sqf_selected)

options(tibble.width = Inf)
```

## Exercise 1

a\. YOUR ANSWER HERE

b\. YOUR ANSWER HERE

c\. YOUR ANSWER HERE

# Summarizing data

## Exercise 2

a\.

```{r}

```

b\. YOUR ANSWER HERE

c\. YOUR ANSWER HERE

d\. YOUR ANSWER HERE

e\. YOUR ANSWER HERE

f\. YOUR ANSWER HERE

g\. YOUR ANSWER HERE

h\. YOUR ANSWER HERE

```{r}
#create new criminal_offense_reported variable
sqf_selected <- sqf_selected |> 
  mutate(criminal_offense_reported = if_else(suspect_arrested_flag == "Y" | summons_issued_flag == "Y", "Y", "N"))
```

## Exercise 3

a\.

```{r}

```

b\. YOUR ANSWER HERE

c\. YOUR ANSWER HERE

d\. YOUR ANSWER HERE

# Comparing across groups

## Exercise 4

a\.

```{r}

```

b\. YOUR ANSWER HERE

c\. YOUR ANSWER HERE

d\. YOUR ANSWER HERE

## Exercise 5

a\.

```{r}

```

b\. YOUR ANSWER HERE

c\. YOUR ANSWER HERE

## Exercise 6

a\.

```{r}

```

b\. YOUR ANSWER HERE

c\. YOUR ANSWER HERE

d\. YOUR ANSWER HERE

e\. YOUR ANSWER HERE

f\. YOUR ANSWER HERE

## Exercise 7

a\.

```{r}
 
```

b\. YOUR ANSWER HERE

c\. YOUR ANSWER HERE

## Exercise 8

YOUR ANSWER HERE

## Exercise 9

a\. YOUR ANSWER HERE

b\. YOUR ANSWER HERE

c\. YOUR ANSWER HERE

d\. YOUR ANSWER HERE

e\. YOUR ANSWER HERE
