---
title: "Lab 05"
subtitle: "Regression"
author: "YOUR NAME HERE"
format:
  html:
    self-contained: true
    toc: true
    toc_float: true
    number_section: false
    highlight: tango
editor: visual
editor_options:
  chunk_output_type: console
---

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(broom)
library(gghighlight)

scorecard <- read_csv("data/scorecard.csv")
```

## Exercise 1

a. YOUR ANSWER HERE

```{r}
#part b code
```


c. YOUR ANSWER HERE

d. YOUR ANSWER HERE

## Exercise 2


```{r}
#part a code
```

b. YOUR ANSWER HERE

## Exercise 3

```{r}
#part a code
```

b. YOUR ANSWER HERE

c. YOUR ANSWER HERE

d. YOUR ANSWER HERE

## Exercise 4

```{r}
#part a code

```

b. YOUR ANSWER HERE

## Exercise 5


```{r}
#part a code
```


b. YOUR ANSWER HERE

c. YOUR ANSWER HERE

d. YOUR ANSWER HERE

```{r}
#part d code
```


```{r}
#part e code
```


```{r}
#part f code
```


g. YOUR ANSWER HERE

## Exercise 6


```{r}
#part a code
```

b. YOUR ANSWER HERE


## Exercise 7

```{r}
#part a code
```


```{r}
#part b code
```


c. YOUR ANSWER HERE

```{r}
#part d code
```

d. YOUR ANSWER HERE 

## Exercise 8

```{r}
#part a-c code

```

d. YOUR ANSWER HERE


## Exercise 9

YOUR ANSWER HERE

### BONUS

a. YOUR ANSWER HERE

```{r}
#part b code
```

c. YOUR ANSWER HERE