---
title: "Exam 2 - Group Takehome"
subtitle: "GROUP # HERE (from Canvas)"
author: "GROUP MEMBER NAMES HERE"
date: "November 13, 2024"
format:
  html:
    self-contained: true
    toc: true
    toc_float: true
    number_section: false
    highlight: tango
editor: visual
editor_options:
  chunk_output_type: console
---

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(infer)
library(kableExtra)

hotels <- read_csv("./data/hotels.csv")
```

## Exercise 1

a\. YOUR ANSWER HERE

b\. YOUR ANSWER HERE

c\. YOUR ANSWER HERE

d\.

-   Numerical discrete:

-   Numerical continuous:

-   Categorical ordinal:

-   Categorical non-ordinal

## Exercise 2

a\. YOUR ANSWER HERE

```{r}
#part b code here
```

c\. YOUR ANSWER HERE

d\. YOUR ANSWER HERE

e\. YOUR ANSWER HERE

f\. YOUR ANSWER HERE

```{r}
#part g code here
```

h\. YOUR ANSWER HERE

i\. YOUR ANSWER HERE

```{r}
#code to create summer variable here
```

## Exercise 3

a\. YOUR ANSWER HERE

```{r}
# part b code here

```

c\. YOUR ANSWER HERE

```{r}
# part d code here
```

e\. YOUR ANSWER HERE

## Exercise 4

```{r}
# part a code 

```

b\. YOUR ANSWER HERE

c\. YOUR ANSWER HERE

```{r}
# part d code here
```

e\. YOUR ANSWER HERE

f\. YOUR ANSWER HERE

g\. YOUR ANSWER HERE

h\. YOUR ANSWER HERE

## Exercise 5

```{r}

```

YOUR ANSWER HERE

## Exercise 6

a\. YOUR ANSWER HERE

```{r}
# part b code 


```

c\. YOUR ANSWER HERE

```{r}
# part d code here
```

e\. YOUR ANSWER HERE

f\. YOUR ANSWER HERE

g\. YOUR ANSWER HERE

## BONUS 1

YOUR ANSWER HERE

```{r}

```

## BONUS 2

a\. YOUR ANSWER HERE

b\. YOUR ANSWER HERE

## 
